---
title: "CD8_dataset_incorporation"
author: "pattiey"
date: '2022-08-23'
output: pdf_document
---

Load packages

```{r results=FALSE, message=FALSE}
library(utils)
library(Seurat)
library(clusterProfiler)
library(dplyr)
library(tidyr)
library(Matrix)
library(sctransform)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(viridis)
library(ComplexHeatmap)
library(circlize)
library(goseq)
library(GEOquery)
library(msigdbr)
library(fgsea)
library(data.table)
library(tibble)
```

Load helper functions

```{r}
source("cd8_analysis.R")
```

Load metadata from GEO

```{r}
B16_GEO <- "GSE116390"
B16_eset <- getGEO(B16_GEO)[[1]]

met_cond_GEO <- "GSE152018"
met_cond_eset <- getGEO(met_cond_GEO)[[1]]
pData(met_cond_eset) <- pData(met_cond_eset) %>% mutate(SampleLabel = ifelse(`treatment:ch1` == "Control (11mM glucose)", "Control", "TGR"))
```

Load aggregated matrix for B16s into Seurat object

```{r}
if (file.exists(file.path("Data", B16_GEO, paste0(B16_GEO, "_aggregated"))) == FALSE) {
  getGEOSuppFiles(B16_GEO, makeDirectory = TRUE, baseDir = "Data", fetch_files = TRUE)
  untar(file.path("Data", B16_GEO, paste0(B16_GEO, "_aggregated_filtered_matrix.tar.gz")), 
      list=FALSE, 
      exdir = file.path("Data", B16_GEO, paste0(B16_GEO, "_aggregated")))
}

B16.counts <- Read10X(file.path("Data", B16_GEO, paste0(B16_GEO, "_aggregated")))
(B16.seurat <- CreateSeuratObject(counts = B16.counts, 
                                   project = "B16CD8",
                                   min.cells = 3,
                                   min.features = 200))
sampleIDs <- qq(PMEL1_S1, WT1_S2, WT3_S3, PMEL3_S1, PMEL2_S1, WT2_S2, WT4_S3)
names(sampleIDs) <- 1:7
B16.seurat[["Sample"]] <- factor(substring(rownames(B16.seurat@meta.data), 18))
B16.seurat[["SampleLabel"]] <- factor(sampleIDs[B16.seurat@meta.data$Sample])
# B16.seurat@meta.data <- left_join(B16.seurat@meta.data, pData(B16_eset), by = c("SampleLabel" = "description"))
```

Load matrix for metabolically conditioned dataset

```{r}
if (file.exists(file.path("Data", met_cond_GEO, paste0(met_cond_GEO, "_RAW"))) == FALSE){
  getGEOSuppFiles(met_cond_GEO, makeDirectory = TRUE, baseDir = "Data", fetch_files = TRUE)
  untar(file.path("Data", met_cond_GEO, paste0(met_cond_GEO, "_RAW.tar")), 
        list = FALSE,
        exdir = met_cond_dir)
}

met_cond_dir <- file.path("Data", met_cond_GEO, paste0(met_cond_GEO, "_RAW"))

dir.create(file.path(met_cond_dir, "CD8_hi"))
dir.create(file.path(met_cond_dir, "CD8_lo"))

file.rename(from = file.path(met_cond_dir, "GSM4598789_CD8_hi_barcodes.tsv.gz"), to = file.path(met_cond_dir, "CD8_hi", "barcodes.tsv.gz"))
file.rename(from = file.path(met_cond_dir, "GSM4598789_CD8_hi_genes.tsv.gz"), to = file.path(met_cond_dir, "CD8_hi", "features.tsv.gz"))
file.rename(from = file.path(met_cond_dir, "GSM4598789_CD8_hi_matrix.mtx.gz"), to = file.path(met_cond_dir, "CD8_hi", "matrix.mtx.gz"))

file.rename(from = file.path(met_cond_dir, "GSM4598790_CD8_lo_barcodes.tsv.gz"), to = file.path(met_cond_dir, "CD8_lo", "barcodes.tsv.gz"))
file.rename(from = file.path(met_cond_dir, "GSM4598790_CD8_lo_genes.tsv.gz"), to = file.path(met_cond_dir, "CD8_lo", "features.tsv.gz"))
file.rename(from = file.path(met_cond_dir, "GSM4598790_CD8_lo_matrix.mtx.gz"), to = file.path(met_cond_dir, "CD8_lo", "matrix.mtx.gz"))
```

```{r}
CD8_hi.counts <- Read10X(file.path(met_cond_dir, "CD8_hi"))
CD8_lo.counts <- Read10X(file.path(met_cond_dir, "CD8_lo"))

(CD8_hi.seurat <- CreateSeuratObject(counts = CD8_hi.counts, 
                                     project = "met_cond",
                                     min.cells = 3,
                                     min.features = 200))

CD8_hi.seurat[["SampleLabel"]] <- "Control"

# CD8_hi.seurat@meta.data <- left_join(CD8_hi.seurat@meta.data, pData(met_cond_eset))

(CD8_lo.seurat <- CreateSeuratObject(counts = CD8_lo.counts, 
                                     project = "met_cond",
                                     min.cells = 3,
                                     min.features = 200))

CD8_lo.seurat[["SampleLabel"]] <- "TGR"

# CD8_lo.seurat@meta.data <- left_join(CD8_lo.seurat@meta.data, pData(met_cond_eset))
```

Normalize and find variable features independently

```{r message=FALSE, results=FALSE, warning=FALSE, include=FALSE}
CD8_hi.seurat <- NormalizeData(CD8_hi.seurat)
CD8_hi.seurat <- FindVariableFeatures(CD8_hi.seurat, selection.method = "vst", nfeatures = 10000)
(top10 <- head(VariableFeatures(CD8_hi.seurat), 10))

CD8_lo.seurat <- NormalizeData(CD8_lo.seurat)
CD8_lo.seurat <- FindVariableFeatures(CD8_lo.seurat, selection.method = "vst", nfeatures = 10000)
(top10 <- head(VariableFeatures(CD8_lo.seurat), 10))
```

integrate CD8_hi and CD8_lo cells

```{r}
data.list <- c(CD8_hi.seurat, CD8_lo.seurat)
features <- SelectIntegrationFeatures(object.list = data.list, nfeatures = 10000)

anchors <- FindIntegrationAnchors(data.list, anchor.features = features)

met_cond.seurat <- IntegrateData(anchorset = anchors)
DefaultAssay(met_cond.seurat) <- "RNA"
```

percentage of ribosomal and mitochondrial genes per cell
```{r}
met_cond.seurat[["percent.mito"]] <- PercentageFeatureSet(met_cond.seurat, pattern = "^mt-")
met_cond.seurat[["percent.ribo"]] <- PercentageFeatureSet(met_cond.seurat, pattern = "^Rp[ls]")
```

```{r}
Idents(met_cond.seurat) <- met_cond.seurat$SampleLabel
my_cols <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#33FF99")
VlnPlot(met_cond.seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"), ncol = 4, pt.size = 0.1, cols = alpha(my_cols, 0.25))
```

```{r}
FeatureScatter(met_cond.seurat, feature1 = "nCount_RNA", feature2 = "percent.mito", cols = alpha(my_cols, 0.25))
```
```{r}
FeatureScatter(met_cond.seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", cols = alpha(my_cols, 0.25))
```
```{r}
FeatureScatter(met_cond.seurat, feature1 = "nCount_RNA", feature2 = "percent.ribo", cols = alpha(my_cols, 0.25))
```
```{r}
FeatureScatter(met_cond.seurat, feature1 = "percent.mito", feature2 = "percent.ribo", cols = alpha(my_cols, 0.25))
```

Filter cells with high mito
```{r}
met_cond.seurat <- subset(met_cond.seurat, subset = percent.mito < 10)
VlnPlot(met_cond.seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo", cols = alpha(my_cols, 0.25)), ncol = 4, pt.size = 0.1)
```

Remove mitochondial and ribosomal genes

```{r}
mito.genes <- grep(pattern = "^mt-", x = rownames(x = met_cond.seurat@assays$RNA@data), value = TRUE)
ribo.genes <- grep(pattern = "^Rp[ls]", x = rownames(x = met_cond.seurat@assays$RNA@data), value = TRUE)
```

```{r}
DefaultAssay(met_cond.seurat) <- "integrated"
met_cond.seurat <- FindVariableFeatures(met_cond.seurat, selection.method = "vst", nfeatures = 7500)
(top10 <- head(VariableFeatures(met_cond.seurat), 10))
length(met_cond.seurat@assays$integrated@var.features)
met_cond.seurat@assays$integrated@var.features <-
  met_cond.seurat@assays$integrated@var.features[!met_cond.seurat@assays$integrated@var.features %in% 
                                        c(mito.genes,ribo.genes)]
length(met_cond.seurat@assays$integrated@var.features)
(top10 <- head(VariableFeatures(met_cond.seurat), 10))
LabelPoints(plot = VariableFeaturePlot(met_cond.seurat), points = top10, repel = TRUE, xnudge = 0, ynudge = 0)
```

Scale data
```{r}
met_cond.seurat <- ScaleData(met_cond.seurat, assay = "integrated")
```

Run PCA and dimensionality reduction

```{r}
set.seed(123)
met_cond.seurat <- dim_reductions(met_cond.seurat)
VizDimLoadings(met_cond.seurat, dims = 1:9, reduction = "pca") + theme(axis.text = element_text(size=5),
                                                                       axis.title = element_text(size=8, face="bold"))

mat <- GetAssayData(met_cond.seurat, assay = "integrated", slot = "scale.data")
pca <- met_cond.seurat[["pca"]]

total_var <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2
(varExplained <- eigValues / total_var)
ElbowPlot(met_cond.seurat)

DimHeatmap(met_cond.seurat, dims = 1:6, nfeatures = 20, cells = 500, balanced = TRUE)
```

```{r}
DimPlot(met_cond.seurat, reduction = "pca", group.by = "SampleLabel", dims = c(1,2), cols = alpha(my_cols, 0.5))
```

UMAP

```{r}
DimPlot(met_cond.seurat, reduction = "umap", group.by = "SampleLabel", dims = c(1,2),  cols = alpha(my_cols, .75))
```
DEGs identified in nature metabolism paper
```{r}
genes_of_interest <- c("Tcf7","Atf4", "Ddit3", "Ifng", "Slc2a1")

FeaturePlot(met_cond.seurat, features = c("2810417H13Rik", "Ranbp1", "Npc2", "Btg1"), reduction = "umap", cols = alpha(c("blue", "red"), 0.5))

FeaturePlot(met_cond.seurat, features = genes_of_interest, reduction = "umap", cols = alpha(c("blue", "red"), 0.25))
met_cond.seurat %>% subset(subset = SampleLabel == "Control") %>% FeaturePlot(features = genes_of_interest, reduction = "umap", cols =alpha(c("blue", "red"), 0.25))
met_cond.seurat %>% subset(subset = SampleLabel == "TGR") %>% FeaturePlot(features = genes_of_interest, reduction = "umap", cols = alpha(c("blue", "red"), 0.25))
```

```{r}
DoHeatmap(met_cond.seurat, features = genes_of_interest, group.by = "SampleLabel", raster=FALSE)
```

# DEA on metabolically conditioned cells

```{r}
TGR.de.markers <- FindMarkers(met_cond.seurat, ident.1 = "Control", ident.2 = "TGR")
TGR.de.markers %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC))) %>% head(20)
TGR.deg <- TGR.de.markers %>% filter(p_val_adj < 0.05)
```

```{r}
FeaturePlot(met_cond.seurat, features = c("Cenpa", "Mt2", "2810417H13Rik"), cols = alpha(c("blue", "red"), 0.25), reduction = "umap")
met_cond.seurat %>% subset(subset = SampleLabel == "Control") %>%  FeaturePlot(features = c("Cenpa", "Mt2", "2810417H13Rik"), cols = alpha(c("blue", "red"), 0.25), reduction = "umap")
met_cond.seurat %>% subset(subset = SampleLabel == "TGR") %>%  FeaturePlot(features = c("Cenpa", "Mt2", "2810417H13Rik"), cols = alpha(c("blue", "red"), 0.25), reduction = "umap")
```

## Hypergeometric GSEA

```{r include=FALSE}
run.hyper.test <- function(gene.dt, geneset.dt, seurat.obj, cutoff = 0.05) {

    .genes <- gene.dt %>%
      rownames_to_column(var = "gene_symbol") %>% 
      select(gene_symbol, p_val_adj)
    .sets <- as.data.table(geneset.dt)
    .sets <- .sets[gene_symbol %in% rownames(seurat.obj),
                   .(gene_symbol, gs_name)]

    .dt <- left_join(.sets, .genes, by = "gene_symbol") %>% 
        as.data.table()

    ## Total number of genes
    ntot <- length(unique(.dt$gene_symbol))
    ## Total number of significant DEGs
    nsig <- nrow(unique(.dt[p_val_adj < cutoff, .(gene_symbol)]))
    ## Gene set size
    gs.size <- .dt[,
                   .(m = length(unique(gene_symbol))),
                   by = .(gs_name)]
    ## Gene set overlap size
    overlap.size <- .dt[p_val_adj < cutoff,
                        .(q = length(unique(gene_symbol))),
                        by = .(gs_name)]

    left_join(gs.size, overlap.size, by = "gs_name") %>%
        mutate(`q` = if_else(is.na(`q`), 0, as.numeric(`q`))) %>% 
        mutate(n = `ntot` - `m`) %>%
        mutate(k = `nsig`) %>%
        mutate(p.val = phyper(`q`, `m`, `n`, `k`, lower.tail=FALSE)) %>%
        arrange(p.val) %>%
        as.data.table
}
```

```{r}
KEGG.mouse.db <- msigdbr(species = "mouse", category = "C2", subcategory = "KEGG")
```

```{r}
hyper.kegg.dt <- run.hyper.test(TGR.deg, KEGG.mouse.db, met_cond.seurat, cutoff = 0.05)
hyper.kegg.dt %>%
  arrange(p.val) %>%
  filter(p.val < 0.05) %>% 
  knitr::kable()
```

## Ranked GSEA

```{r}
num.sci <- function(x) {
    format(x, scientific=TRUE, digits = 2)
}

make.gs.lol <- function(.dt) {
  .dt <- as.data.table(.dt) %>% unique()
  .list <- 
    .dt[, .(gene = .(gene_symbol)), by = .(gs_name)] %>% 
    as.list()
  .names <- .list$gs_name
  .ret <- .list$gene
  names(.ret) <- .names
  return(.ret)
}

KEGG.lol <- KEGG.mouse.db %>% select(gene_symbol, gs_name) %>% make.gs.lol()
```

```{r}
ggplot(TGR.de.markers, aes(1:nrow(TGR.de.markers), -log10(p_val_adj))) + 
  geom_point(stroke = 0) + 
  geom_hline(yintercept = 4, lty = 2, colour = "red") + 
  geom_hline(yintercept = 32, lty = 2, colour = "green") +
  geom_hline(yintercept = 64, lty = 2, colour = "blue") + 
  scale_y_continuous("adjusted p-val", breaks = c(4,32,64), 
                     labels = function(x) num.sci(10^(-x)))
```

```{r}
deg.scores <- TGR.de.markers %>%
  mutate(adj_p_val = p_val_adj + 1e-300) %>% 
  mutate(v = -log10(adj_p_val)) %>% 
  (function(.dt) {v <- .dt$v; names(v) <- rownames(.dt); v})
```

```{r}
kegg.fgsea <- fgsea(pathways = KEGG.lol, stats = deg.scores, scoreType = "pos")
kegg.fgsea[,
           topGenes := paste0(head(unlist(`leadingEdge`), 3), collapse=", "),
           by = .(pathway)]

kegg.fgsea %>%
  arrange(pval) %>%
  filter(pval < 0.1) %>% 
  select(-leadingEdge) %>% 
  knitr::kable()
```

QC and filtering on B16 dataset
```{r}
B16.seurat[["percent.mito"]] <- PercentageFeatureSet(B16.seurat, pattern = "^mt-")
B16.seurat[["percent.ribo"]] <- PercentageFeatureSet(B16.seurat, pattern = "^Rp[ls]")
VlnPlot(B16.seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"), ncol = 4, pt.size = 0.1)
```

Normalize and scale
```{r}
B16.seurat <- NormalizeData(B16.seurat)
B16.seurat <- FindVariableFeatures(B16.seurat, selection.method = "vst", mean.cutoff = c(0.0125, 3), dispersion.cutoff = c(0.5, Inf))
(top10 <- head(VariableFeatures(B16.seurat), 10))
```

```{r}
length(B16.seurat@assays$RNA@var.features)
B16.seurat@assays$RNA@var.features <-
  B16.seurat@assays$RNA@var.features[!B16.seurat@assays$RNA@var.features %in% 
                                        c(mito.genes,ribo.genes)]
length(B16.seurat@assays$RNA@var.features)
(top10 <- head(VariableFeatures(B16.seurat), 10))
LabelPoints(plot = VariableFeaturePlot(B16.seurat), points = top10, repel = TRUE, xnudge = 0, ynudge = 0)

B16.seurat <- ScaleData(B16.seurat, assay = "RNA")
```

```{r}
set.seed(123)
B16.seurat <- dim_reductions(B16.seurat)
VizDimLoadings(B16.seurat, dims = 1:9, reduction = "pca") + theme(axis.text = element_text(size=5),
                                                                       axis.title = element_text(size=8, face="bold"))

mat <- GetAssayData(B16.seurat, assay = "RNA", slot = "scale.data")
pca <- B16.seurat[["pca"]]

total_var <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2
(varExplained <- eigValues / total_var)
ElbowPlot(B16.seurat)

DimHeatmap(B16.seurat, dims = 1:6, nfeatures = 20, cells = 500, balanced = TRUE)
```

```{r}
DimPlot(B16.seurat, reduction = "pca", group.by = "SampleLabel", dims = c(1,2), cols = alpha(my_cols, 0.5))
```

```{r}
DimPlot(B16.seurat, reduction = "umap", group.by = "SampleLabel", dims = c(1,2), cols = alpha(my_cols, 0.5))
```

# Combine datasets

```{r}
data.list <- c(CD8_hi.seurat, CD8_lo.seurat, B16.seurat)
features <- SelectIntegrationFeatures(object.list = data.list, nfeatures = 10000)

anchors <- FindIntegrationAnchors(data.list, anchor.features = features)

combined <- IntegrateData(anchorset = anchors)
DefaultAssay(combined) <- "RNA"
```

Let's just see what happens?
```{r}
combined[["percent.mito"]] <- PercentageFeatureSet(combined, pattern = "^mt-")
combined[["percent.ribo"]] <- PercentageFeatureSet(combined, pattern = "^Rp[ls]")
```

```{r}
VlnPlot(combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"), ncol = 4, pt.size = 0.1, cols = alpha(my_cols, 0.5))
```

```{r}
FeatureScatter(combined, feature1 = "nCount_RNA", feature2 = "percent.mito", cols = alpha(my_cols, 0.5))
```
```{r}
FeatureScatter(combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", cols = alpha(my_cols, 0.5))
```
```{r}
FeatureScatter(combined, feature1 = "nCount_RNA", feature2 = "percent.ribo", cols = alpha(my_cols, 0.5))
```
```{r}
FeatureScatter(combined, feature1 = "percent.mito", feature2 = "percent.ribo", cols = alpha(my_cols, 0.5))
```

Filter cells with high mito
```{r}
combined <- subset(combined, subset = percent.mito < 10)
VlnPlot(combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"), ncol = 4, pt.size = 0.1, cols = alpha(my_cols, 0.5))
```

Remove mitochondial and ribosomal genes

```{r}
mito.genes <- grep(pattern = "^mt-", x = rownames(x = combined@assays$RNA@data), value = TRUE)
ribo.genes <- grep(pattern = "^Rp[ls]", x = rownames(x = combined@assays$RNA@data), value = TRUE)
```

```{r}
DefaultAssay(combined) <- "integrated"
combined <- FindVariableFeatures(combined, selection.method = "vst", nfeatures = 7500)
(top10 <- head(VariableFeatures(combined), 10))
length(combined@assays$integrated@var.features)
combined@assays$integrated@var.features <-
  combined@assays$integrated@var.features[!combined@assays$integrated@var.features %in% 
                                        c(mito.genes,ribo.genes)]
length(combined@assays$integrated@var.features)
(top10 <- head(VariableFeatures(combined), 10))
LabelPoints(plot = VariableFeaturePlot(combined), points = top10, repel = TRUE, xnudge = 0, ynudge = 0)
```

Scale data
```{r}
combined <- ScaleData(combined, assay = "integrated")
```

Run PCA and dimensionality reduction

```{r}
set.seed(123)
combined <- dim_reductions(combined)
VizDimLoadings(combined, dims = 1:9, reduction = "pca") + theme(axis.text = element_text(size=5),
                                                                axis.title = element_text(size=8, face = "bold"))

mat <- GetAssayData(combined, assay = "integrated", slot = "scale.data")
pca <- combined[["pca"]]

total_var <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2
(varExplained <- eigValues / total_var)
ElbowPlot(combined)

DimHeatmap(combined, dims = 1:6, nfeatures = 20, cells = 500, balanced = TRUE)
```

```{r}
combined$SampleLabel <- factor(as.factor(combined$SampleLabel), levels = c("Control", "TGR", "PMEL1_S1", "PMEL2_S1", "PMEL3_S1", "WT1_S2", "WT2_S2", "WT3_S3", "WT4_S3"))
DimPlot(combined, reduction = "pca", group.by = "SampleLabel", dims = c(1,2), cols = alpha(my_cols, 0.75))
DimPlot(combined, reduction = "pca", dims = c(1,2), group.by = "orig.ident", cols = alpha(my_cols, 0.75))
combined %>% subset(orig.ident == "met_cond") %>% DimPlot(reduction = "pca", group.by = "SampleLabel", cols = alpha(my_cols, 0.75))
```

UMAP

```{r}
DimPlot(combined, reduction = "umap", group.by = "SampleLabel", dims = c(1,2), cols = alpha(my_cols, 0.75))
DimPlot(combined, reduction = "umap",  group.by = "orig.ident", dims = c(1,2), cols = alpha(my_cols, 0.75))
```

incorporate clusters
```{r}
pre.clusters <- read.csv(file.path("Data", B16_GEO, "cellClusters.csv"))
pre.clusters.meta <- pre.clusters[match(substring(colnames(combined), 1, 18),pre.clusters$X),2]
pre.clusters.meta <- factor(pre.clusters.meta, levels = c("Control", "TGR", "Naive", "EM-like", "MemoryLike", "Exhausted"))
pre.clusters.meta[is.na(pre.clusters.meta)] <- combined@meta.data$SampleLabel[is.na(pre.clusters.meta)]
names(pre.clusters.meta) <- colnames(combined)
combined <- AddMetaData(combined,pre.clusters.meta,"cluster_label")
```

```{r}
FeaturePlot(combined,features=genes_of_interest,reduction="umap", cols = alpha(c("blue", "red"), 0.25)) 
```

```{r}
DimPlot(combined, group.by = "cluster_label", reduction = "pca", cols = alpha(my_cols, 0.5))
DimPlot(combined, group.by = "cluster_label", reduction = "umap", cols = alpha(my_cols, 0.5))
```

```{r}
combined %>% subset(subset = cluster_label != "TGR" & cluster_label != "Control") %>% DimPlot(group.by = "cluster_label", reduction = "umap", cols = alpha(my_cols, 0.5))
combined %>% subset(subset = cluster_label != "Control") %>% DimPlot(group.by = "cluster_label", reduction = "umap", cols = alpha(my_cols, 0.5))
combined %>% subset(subset = cluster_label != "TGR") %>% DimPlot(group.by = "cluster_label", reduction = "umap", cols = alpha(my_cols, 0.5))
```

```{r}
DoHeatmap(combined, features = genes_of_interest, group.by = "cluster_label", raster=FALSE)
```

# Cluster

```{r}
combined <- FindNeighbors(combined, reduction = "pca", dims = 1:10, k.param = 30)
combined <- FindClusters(combined, resolution = 0.5)
```

```{r}
DimPlot(combined, group.by = "seurat_clusters", reduction = "umap", cols = alpha(my_cols, 0.5))
DimPlot(combined, group.by = "cluster_label", reduction = "umap", cols = alpha(my_cols, 0.5))
```

```{r}
DoHeatmap(combined, features = genes_of_interest, group.by = c("seurat_clusters", "cluster_label"), raster=FALSE)
```

```{r}
prop.table(table(combined$seurat_clusters, combined$cluster_label), margin = 1) %>% as.data.frame() %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + xlab("Seurat Cluster") + labs(fill = "Sample Label")
prop.table(table(combined$seurat_clusters, combined$cluster_label), margin = 2) %>% as.data.frame() %>% ggplot(aes(x = Var2, y = Freq, fill = Var1)) + geom_bar(stat = "identity") + labs(x = "Sample Label", fill = "Seurat Cluster")

table(combined$seurat_clusters, combined$cluster_label) %>% as.data.frame() %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + xlab("Seurat Cluster") + labs(fill = "Sample Label")
table(combined$seurat_clusters, combined$cluster_label) %>% as.data.frame() %>% ggplot(aes(x = Var2, y = Freq, fill = Var1)) + geom_bar(stat = "identity") + xlab("Sample Label") + labs(fill = "Seurat Cluster")
```

# Cluster 2 DEA and GSEA

```{r}
cluster.2.markers <- FindMarkers(combined, ident.1 = 2)
cluster.2.markers %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC))) %>% head(20)
cluster.2.deg <- cluster.2.markers %>% filter(p_val_adj < 0.05)
```

## Hypergeometric GSEA

```{r}
hyper.kegg.dt <- run.hyper.test(cluster.2.deg, KEGG.mouse.db, combined, cutoff = 0.05)
hyper.kegg.dt %>%
  arrange(p.val) %>%
  filter(p.val < 0.05) %>% 
  knitr::kable()
```

```{r}
ggplot(cluster.2.markers, aes(1:nrow(cluster.2.markers), -log10(p_val_adj))) + 
  geom_point(stroke = 0) + 
  geom_hline(yintercept = 4, lty = 2, colour = "red") + 
  geom_hline(yintercept = 32, lty = 2, colour = "green") +
  geom_hline(yintercept = 64, lty = 2, colour = "blue") + 
  scale_y_continuous("adjusted p-val", breaks = c(4,32,64), 
                     labels = function(x) num.sci(10^(-x)))
```

```{r}
deg.scores <- cluster.2.deg %>%
  mutate(adj_p_val = p_val_adj + 1e-300) %>% 
  mutate(v = -log10(adj_p_val)) %>% 
  (function(.dt) {v <- .dt$v; names(v) <- rownames(.dt); v})
```

## Ranked GSEA

```{r}
kegg.fgsea <- fgsea(pathways = KEGG.lol, stats = deg.scores, scoreType = "pos")
kegg.fgsea[,
           topGenes := paste0(head(unlist(`leadingEdge`), 3), collapse=", "),
           by = .(pathway)]

kegg.fgsea %>%
  arrange(pval) %>%
  filter(pval < 0.1) %>% 
  select(-leadingEdge) %>% 
  knitr::kable()
```


