---
title: "CD8_analysis2"
author: "pattiey"
date: "21/10/2021"
output: pdf_document
---

Analysis of CD8+ T cells. Much of this is borrowed from the workflow found in the paper of the original dataset (Carmona et al.) found here https://gitlab.unil.ch/carmona/workflow_Carmona_etal_2019_CD8TIL_Sv3/blob/master/workflow1.Rmd 

```{r results=FALSE, message=FALSE}
library(Seurat)
library(clusterProfiler)
library(dplyr)
library(tidyr)
library(Matrix)
library(sctransform)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(viridis)
library(ComplexHeatmap)
library(circlize)
library(goseq)
```

```{r}
source("cd8_analysis.R")
```

Read sparse matrix

```{r}
fname <- "input/B16_data"
data <- Read10X(fname)
(data.seurat <- CreateSeuratObject(counts = data, 
                                   project = "B16CD8",
                                   min.cells = 3, 
                                   min.features = 50))
```

Determine proportion of mitochondrial and ribosomal genes
```{r}
data.seurat[["percent.mito"]] <- PercentageFeatureSet(data.seurat, pattern = "^mt-")
data.seurat[["percent.ribo"]] <- PercentageFeatureSet(data.seurat, pattern = "^Rp[ls]")
```

```{r}
sampleIDs <- qq(PMEL1, WT1, WT3, PMEL3, PMEL2, WT2, WT4)
names(sampleIDs) <- 1:7
```

```{r}
data.seurat[["Sample"]] <- factor(substring(rownames(data.seurat@meta.data), 18))
data.seurat[["SampleLabel"]] <- factor(sampleIDs[data.seurat@meta.data$Sample])
```

Normalize UMI count data with SCTransform

```{r message=FALSE, results=FALSE, warning=FALSE, include=FALSE}
data.seurat <- SCTransform(data.seurat, variable.features.n=10000, verbose = FALSE)
```

Run PCA 

```{r}
set.seed(123)
DefaultAssay(object = data.seurat) <- "SCT"
data.seurat <- RunPCA(data.seurat, ndims.print = 1:5, nfeatures.print = 10)
```

Remove cell cycle (excluding our genes of interest), ribosomal, and mitochondrial genes from genes used for dimensionality reduction.
```{r}
cellCycle.genes <- read.csv("cellCycleGenes.csv", as.is = T)$x
mito.genes <- grep(pattern = "^mt-", x = rownames(x = data.seurat@assays$RNA@data), value = TRUE)
ribo.genes <- grep(pattern = "^Rp[ls]", x = rownames(x = data.seurat@assays$RNA@data), value = TRUE)
```

```{r}
length(data.seurat@assays$SCT@var.features)
data.seurat@assays$SCT@var.features <-
  data.seurat@assays$SCT@var.features[!data.seurat@assays$SCT@var.features %in% 
                                        c(mito.genes,ribo.genes, cellCycle.genes)]
length(data.seurat@assays$SCT@var.features)
```

Run dimensionality reduction with mitochondrial and ribosomal genes removed

```{r}
data.seurat <- dim_reductions(data.seurat, seed = 123)
```

PCA coloured by the 5 genes of interest

```{r}
genes_of_interest <- c("Tcf7","Atf4", "Ddit3", "Ifng", "Slc2a1")
FeaturePlot(data.seurat,features=genes_of_interest,reduction="pca") 
FeaturePlot(data.seurat, features=genes_of_interest, dims = c(5,4), reduction = "pca")
```

Use KNN to find neighbours on PCA dimensions

```{r}
set.seed(123)
data.seurat <- FindNeighbors(data.seurat, reduction = "pca", dims = 1:10,k.param = 30)
set.seed(123)
data.seurat  <- FindClusters(data.seurat, resolution = 0.3)
```

Add original inferred cell state clustering to meta data
```{r}
pre.clusters <- read.csv("cellClusters.csv")
pre.clusters.meta <- pre.clusters[match(colnames(data.seurat),pre.clusters$X),2]
names(pre.clusters.meta) <- colnames(data.seurat)
pre.clusters.meta <- factor(pre.clusters.meta,levels=c("Naive","EM-like","MemoryLike","Exhausted"))
data.seurat <- AddMetaData(data.seurat,pre.clusters.meta,"original_clusters")
```

UMAP plot coloured by pca clusters and the original clusters
```{r}
clusterColors <- c("#F8766D","#A58AFF","#00B6EB","#53B400","#E8E8E8")
names(clusterColors) <- c("Naive","EM-like","MemoryLike","Exhausted","Cycling") 
side_by_side_cluster_plot(data.seurat, 
                          reduction = "umap", 
                          group.by1 = "original_clusters", 
                          group.by2 = "seurat_clusters",
                          clusterColors = clusterColors)
```

tSNE plot coloured by pca clusters and the original clusters

```{r}
side_by_side_cluster_plot(data.seurat, 
                          reduction = "tsne", 
                          group.by1 = "original_clusters", 
                          group.by2 = "seurat_clusters", 
                          clusterColors = clusterColors)
```

Cluster based on UMAP dimensions

```{r}
set.seed(123)
data.seurat <- FindNeighbors(data.seurat, reduction = "umap", dims = 1:2, k.param = 30)
set.seed(123)
data.seurat  <- FindClusters(data.seurat, resolution = 0.08)
```

UMAP plot coloured by UMAP clusters and the original clusters
```{r}
side_by_side_cluster_plot(data.seurat, 
                          reduction = "umap", 
                          group.by1 = "original_clusters",
                          group.by2 = "seurat_clusters", 
                          clusterColors = clusterColors)
```

tSNE plot coloured by UMAP clusters and the original clusters
```{r}
side_by_side_cluster_plot(data.seurat, 
                          reduction = "tsne", 
                          group.by1 = "original_clusters",
                          group.by2 = "seurat_clusters", 
                          clusterColors = clusterColors)
```

```{r}
FeaturePlot(data.seurat,features=genes_of_interest,reduction="tsne") 
```

```{r}
FeaturePlot(data.seurat,features=genes_of_interest,reduction="umap") 
```

```{r}
DoHeatmap(data.seurat, features = genes_of_interest, group.by = "original_clusters", raster = FALSE) + scale_fill_viridis(option = "magma")
```

```{r}
DoHeatmap(data.seurat, features = genes_of_interest, group.by = "seurat_clusters", raster = FALSE) + scale_fill_viridis(option = "magma")
```

Now look only at subsets of cells by inferred cell state. To maintain consistency, I'll use the inferred cell states given by `original_clusters`, even if the UMAP clustering may be more accurate. 

```{r}
Idents(data.seurat) <- "original_clusters"
```

### Naive

```{r warning=FALSE}
naive <- subset_setup(data.seurat, idents = "Naive")
naive <- dim_reductions(naive, seed = 123)
```

```{r}
FeaturePlot(naive, features = genes_of_interest, reduction = "pca") 
```

```{r}
FeaturePlot(naive, features = genes_of_interest, reduction = "umap") 
```

```{r warning=FALSE}
naive_correlations <- subset_correlations(naive, genes_of_interest)
```

```{r}
naive_highly_correlated <- subset_highly_correlated(naive_correlations, genes_of_interest, n_to_pull = 100)
Heatmap(naive_highly_correlated, show_row_names = FALSE)
heatmap_correlated_genes(naive, naive_highly_correlated)
```

### EM-like

```{r warning=FALSE}
em_like <- subset_setup(data.seurat, idents = "EM-like")
em_like <- dim_reductions(em_like, seed = 123)
```

```{r}
FeaturePlot(em_like, features = genes_of_interest, reduction = "pca") 
```

```{r}
FeaturePlot(em_like, features = genes_of_interest, reduction = "umap") 
```

```{r warning=FALSE}
em_correlations <- subset_correlations(em_like, genes_of_interest)
```

```{r}
em_highly_correlated <- subset_highly_correlated(em_correlations, genes_of_interest, n_to_pull = 100)
Heatmap(em_highly_correlated)
heatmap_correlated_genes(em_like, em_highly_correlated)
```

### Memory-like

```{r warning=FALSE}
memory_like <- subset_setup(data.seurat, idents = "MemoryLike")
memory_like <- dim_reductions(memory_like, seed = 123)
```

```{r}
FeaturePlot(memory_like, features = genes_of_interest, reduction = "pca") 
```

```{r}
FeaturePlot(memory_like, features = genes_of_interest, reduction = "umap") 
```

```{r warning=FALSE}
memory_correlations <- subset_correlations(memory_like, genes_of_interest)
```

```{r}
memory_highly_correlated <- subset_highly_correlated(memory_correlations, genes_of_interest, n_to_pull = 100)
Heatmap(memory_highly_correlated)
heatmap_correlated_genes(memory_like, memory_highly_correlated)
```

### Exhausted

```{r warning=FALSE}
exhausted <- subset_setup(data.seurat, idents = "Exhausted")
exhausted <- dim_reductions(exhausted, seed = 123)
```

```{r}
FeaturePlot(exhausted, features = genes_of_interest, reduction = "pca") 
```

```{r}
FeaturePlot(exhausted, features = genes_of_interest, reduction = "umap") 
```

```{r warning=FALSE}
exhausted_correlations <- subset_correlations(exhausted, genes_of_interest)
```

```{r}
exhausted_highly_correlated <- subset_highly_correlated(exhausted_correlations, genes_of_interest, n_to_pull = 100)
Heatmap(exhausted_highly_correlated)
heatmap_correlated_genes(exhausted, exhausted_highly_correlated)
```


## Alternative clustering?

See if any different patterns emerge if UMAP clustering is used over the original given clustering.

```{r}
Idents(data.seurat) <- "seurat_clusters"
```

### Naive

```{r}
umap_naive <- subset(data.seurat, idents = 1)
```

### EM like

```{r}
umap_em <- subset(data.seurat, idents = 0)
```

### Memory like

```{r}
umap_memory <- subset(data.seurat, idents = 3)
```

### Exhausted

```{r}
umap_exhausted <- subset(data.seurat, idents = 2)
```
